Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} model system

### 2014-04-25 paper_enahpc14/Sec4:Experimental-results.tex
WS: proofread

For all the experiments we also analyze the contribution of the MPI library to the energy consumption and how the use of the blocking versus polling message-passing policies can potentially render energy savings. Specifically, the OpenMPI 1.6.5 library installed in \tinto, features two operation modes, blocking and modes, which can be selected before \cosmoart is launched. In the polling configuration (the detault mode) ``MPI'' continually polls
the netwrok interface to check for the completion of an event (e.g. send or receive).
This usually yields low latency but high CPU utilization. Thus, can one expcet that this mode attanins the best performance, possibley at the cost a higher energy usage. In the blocking
mode, the CPU is yield to other processes/threads if there are no incoming messages.

@@

Figures~\ref{fig:1}   and   \ref{fig:2}   account   respectively   for
\monch's Isola E1 Rack  2 and \pilat' Isola HD total
power measurements for  1-day or 2-days simulations. On  the Intel Ivy
Bridge EP  based cluster (i.e.  \monch),  the 1-day simulation
was issued  only twice due  to usage restrictions. As  time resolution
was set to one update every  5 minutes for power sampling, the average
power consumption was computed by considering 6 values for each single
run.  On the Intel Xeon E5 based cluster (i.e.  \pilat), the
1-day simulation  was issued  four times and  a 2-days run  only once.
Similarly, the average power consumption was computed by considering 4
values  for  each  single  1-day  run  and 9  values  for  the  2-days
run. Corresponding results are gathered in Table~\ref{tab:3}.

@@

\begin{figure}[htbf]
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-5.eps}
  \caption{Mean time-to-solution and energy-to-solution on
    \textsc{Tintorrum} for a 24h simulation using the polling and blocking MPI policies.}
  \label{fig:4}
\end{figure}

@@

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution. First we run \cosmoart for a full-day simulation
and analyze the power consumption. Second we correlate performance with
power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23h-24h). All the experiments were performed using 192 \cosmoart processes
mapping 12 processes in each of node.

For all the experiments we also analyze the contribution of the MPI library to the energy 
consumption and how the use of the blocking versus polling message-passing policies can 
potentially render energy savings. Specifically, the OpenMPI 1.6.5 library installed in \tinto, 
features two operation modes, blocking and modes, which can be selected before \cosmoart is 
launched. In the polling configuration (the detault mode) ``MPI'' continually polls
the netwrok interface to check for the completion of an event (e.g. send or receive).
This usually yields low latency but high CPU utilization. Thus, can one expcet that this mode attanins the best performance, possibley at the cost a higher energy usage. In the blocking
mode, the CPU is yield to other processes/threads if there are no incoming messages.

In order to obtain an initial overview of the power dissipation of the system model, we have 
obtained the power profiles using our \pmlib framework of a full-day simulation of \cosmoart 
leveraging the polling and blocking MPI policies, these results are depicted in the top/bottom 
plots of Figure~\ref{fig:9}, respectvely. o, captured with our PDUs. The first observation for 
the polling MPI policy is the quite plain power profile it generates. However, the first and 
last hours of the full-day simulation dissipate slightly less power than the mid-day hours. We 
relate these variations due to the chemical reactions and aerosols, computed by ART, that are 
taking place during the daylight hours but not during night. In this case peak power 
dissipation is about 3821\,W. Looking at the power profile using the blocking MPI policy, we 
can easily observe that the pattern is repeated favoring the reduction of power during 
nighttime hours in a more  noticeable way. The peak power dissipation in this mode is 3768\,W, 
which means a reduction of 50\,W in total or, in other words, 1,\% of reductuon in the peak 
power. Also, we observe power  drops in each hour simulated, basically related to periods in 
which the processes were performing blocking-waits instead of polling-waits, thus yielding the 
CPU to a low usage during this phases. Our final observation is the reduction in execution time 
when the blocking MPI policy is leveraged in contrast with the polling MPI. One can expect that 
the costs of the interruptions used in the blocking can lead to an increase of the total 
execution time, however, applications with extensive computing may show better perforamnce with 
the blocking mode. This plays in favour for our results since the energy consumed is also less 
by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

>>>>

For the experiments with \tinto platform, we also analyse the contribution of the MPI library to the energy consumption and how the use of the blocking versus polling message-passing policies can potentially render energy savings. Specifically, the OpenMPI 1.6.5 library, 
features two operation modes, blocking and polling, which can be selected before \cosmoart
is launched. In the polling configuration (the detault mode) ``MPI'' continually polls
the network interface to check for the completion of an event (e.g. send or receive).
This usually yields low latency but high CPU utilization. Thus, one can expcet that this mode 
attains the best performance, possibly at the cost a higher energy usage. In the blocking
mode, the CPU is yield to other processes/threads if there are no incoming messages.

@@

Figures~\ref{fig:1}   and   \ref{fig:2}   account   respectively   for
\monch's Isola E1 Rack  2 and \pilat' Isola HD total
power measurements for  1-day or 2-day simulations. On  the Intel Ivy
Bridge EP  based cluster (i.e.  \monch),  the 1-day simulation
was issued  only twice due  to usage restrictions. As  time resolution
was set to one update every  5 minutes for power sampling, the average
power consumption was computed by considering 6 values for each single
run.  On the Intel Xeon E5 based cluster (i.e.  \pilat), the
1-day simulation  was issued  four times and  a 2-day run  only once.
Similarly, the average power consumption was computed by considering 4
values  for  each  single  1-day  run  and 9  values  for  the  2-days
run. Corresponding results are gathered in Table~\ref{tab:3}.

@@

\begin{figure}[htbf]
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-5.eps}
  \caption{Mean time-to-solution and energy-to-solution on
    \textsc{Tintorrum} for a 24h simulation using the polling and blocking MPI policies. The 95\% 
            confidence intervals of Student's $t$ distribution (10 samples) clearly indicates 
            an improvement for blocking policy in both TTS and ETS.}
  \label{fig:4}
\end{figure}

@@

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution using the attached PDUs in each node. First we run \cosmoart 
for a full-day simulation and analyse the power consumption. Second we correlate performance 
with power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23h-24h). All the experiments were performed using 192 \cosmoart processes
mapping 12 processes in each of node.

% For all the experiments we also analyse the contribution of the MPI library to the energy 
% consumption and how the use of the blocking versus polling message-passing policies can 
% potentially render energy savings. Specifically, the OpenMPI 1.6.5 library installed in \tinto, 
% features two operation modes, blocking and modes, which can be selected before \cosmoart is 
% launched. In the polling configuration (the detault mode) ``MPI'' continually polls
% the netwrok interface to check for the completion of an event (e.g. send or receive).
% This usually yields low latency but high CPU utilization. Thus, can one expcet that this mode attanins the best performance, possibley at the cost a higher energy usage. In the blocking
% mode, the CPU is yield to other processes/threads if there are no incoming messages.

In order to obtain an initial overview of the power dissipation of the system model, we have 
obtained the power profiles using our \pmlib framework of a full-day simulation of \cosmoart 
leveraging the polling and blocking MPI policies, these results are depicted in the top/bottom 
plots of Figure~\ref{fig:9}, respectvely. The first observation for 
the polling MPI policy is the quite plain power profile it generates. However, the first and 
last hours of the full-day simulation dissipate slightly less power than the mid-day hours. We 
relate these variations due to the chemical reactions and aerosols, computed by ART, that are 
taking place during the daylight hours but not during night. In this case power consumption
 is about 3821\,W. Looking at the power profile using the blocking MPI policy, we 
can easily observe that the pattern is repeated favoring the reduction of power during 
nighttime hours in a more  noticeable way. The peak power dissipation in this mode is 3768\,W, 
which means a reduction of 50\,W in total or, in other words, 1\,\% of reductuon in the peak 
power. Also, we observe power  drops in each hour simulated, basically related to periods in 
which the processes were performing blocking-waits instead of polling-waits, thus yielding the 
CPU to a low usage during this phases. Our final observation is the reduction in execution time 
when the blocking MPI policy is leveraged in contrast with the polling MPI. One can expect that 
the costs of the interruptions used in the blocking can lead to an increase of the total 
execution time, however, applications with extensive computing may show better perforamnce with 
the blocking mode. This plays in favour for our results since the energy consumed is also less 
by just changing the behavior of the OpenMPI library from the polling to the blocking policy.
